Learning constraints in tabular data

### 2016-05-20 Paper/ECAI/tabular.tex
fixing things catching my eye

The question that we answer in this paper is: is it possible to discover or reconstruct structural constraints (relations, functions) in flat tabular spreadsheet data?
To answer this question we contribute a general-purpose method and system, named \sname (pronounced as the word ``tackle''), for discovering row-wise and column-wise constraints.
%Furthermore, the algorithm accepts declarative specifications of the constraints and predicates of interest (the "bias"). \tias{revise declarative claim?}
It operates directly on (headerless) tables of a spreadsheet in an unsupervised setting, as it reasons on that raw tabular data directly with no example constraint instantiations given. %, and discovers column- and row-wise constraints.
We demonstrate the utility of our approach in an experimental evaluation.
Moreover, we sketch additional application scenario's such as autocompletion and error detection.
\\\\
This paper is organized as follows.
Section \ref{sec:formalization} introduces concepts relevant to our approach. Section \ref{sec:problem_statement} presents the problem statement and the approach. Section \ref{sec:evaluation} presents the evaluation of the approach. Section \ref{sec:applications} shows how our system can be used for applications. Section \ref{sec:related_work} presents an overview of the related work. Section \ref{sec:conclusions} provides conclusions.

@@

Our proposed methodology contains the following steps:
\begin{enumerate}
\item Extract headerless tables from tabular data
\item Partition the tables into maximally contiguous, type-consistent \textit{superblocks}
\item Generate for each constraint all valid argument assignments:
\begin{enumerate}
\item For each constraint $c$, generate all atoms $c(B_1, \ldots ,B_n)$ where each $B_i$ is a maximal \textit{superblock} and the $B_i$ are compatible with the signature.
\item For each generated atom $c(B_1, \ldots ,B_n)$, find all valid $c(B'_1, \ldots, B'_n)$ such that for all $i$ holds $B'_i \sqsubseteq B_i$ and the signature and definition of $c(B'_1, \ldots, B'_n)$ are satisfied.
\end{enumerate}
\end{enumerate}

@@

\begin{figure}[t]
  \centering
  \includegraphics[width=1\linewidth]{figures/tabletool.png}
  \caption{This figure shows our table selection tool. A user selects rectangular ranges of cells that correspond to (headerless) tables. If the data has an unambiguous orientation, this information can be provided as well.}
  \label{fig:visual_tool}
\end{figure}

@@

\paragraph{Example}
As mentioned before, $\eclookup{G'_2}{G'_1}{H'_1}{H'_2}$ implies $\ecfkey{G'_1}{H'_1}$ and hence the $LOOKUP$ constraint template depends on $FOREIGNKEY$ in Figure~\ref{fig:learning_order} and will be earlier in the ordering.

@@

Our implementation supports the ASP \cite{whaisasp} formalism, using the stand-alone Clingo 4.5.4 \cite{clingo} solver, the stand-alone Constraint Programming language MiniZinc \cite{minizinc} with the Gecode 4.4.0 solver which supports floats, as well as an embedded Python-based CP solver \cite{python_constraint}. The constraint signature that must be satisfied to generate the groups can be specified in each of the tree formalisms, to ease prototyping and extendability towards new constraints. The embedded solver is fastest as it does not have the overhead of file-writing and calling an external program, and is used in the experiments below.

@@

Extending the system with a new constraint template amounts to \tias{TODO: samuel, explain how to define syntax; signature-for-groupgen; signature-for-def; definition}

\sergey{under question now, might worth deleting} \\
\sergey{example new template: vr = floor(v), float(v), int(vr), len(v)=len(vr), v,vr are single vectors?}

\begin{equation}
  \forall i{:}~ v_r[i] \leq v[i] < v_r[i] + 0.5
  \label{eq:floor_extension}
\end{equation}

@@

\subsubsection{Results}
Our current implementation takes a few seconds to find 18 constraints, including all of the 12 solutions described in Figure~\ref{fig:sol_example}.
Of the 6 remaining (redundant) constraints, 5 are $\mathit{rank}$ constraints that are true by accident, such as: \begin{align*}
  & \ecrank{\range{T_1}{\rangeall}{1}}{\range{T_1}{\rangeall}{5}}
\end{align*}
The other redundant constraint is an additional $\mathit{lookup}$ that holds because the two vectors in \range{T_2}{\rangeall}{\rangeto{2}{3}} can both be used to look each other up in \range{T_1}{\rangeall}{\rangeto{1}{2}} and we considered only one of them to be essential (lookup using the ID).

@@

\subsection{Benchmark}
There are three main categories of spreadsheets in the benchmark: spreadsheets from an exercise session teaching Excel at an affiliated University, spreadsheets from tutorials online and publicly available spreadsheets such as crime statistics or financial reports that demonstrate more real world usage\footnote{(attached to to the submitted paperthe submitted paper)}.
The case study is also included.

@@

\subsection{Autocompletion}
\begin{figure}[thb]
  \begin{center}
    \includegraphics[width=0.33\textwidth]{figures/autocompletion_example.png}
  \end{center}
  \caption{\sergey{Samuel, Tias have a look at it, should be corrected and extended, like with groups indications and formulas?}}
  \label{fig:autocompletion_example}
\end{figure}

@@

\sergey{there should be a discussion on what to do if multiple constraints satisfy the data and propose predictions, just a sketch here for now}
If multiple constraints satisfy the data and allow autocompletion, then there is a number of ways to handle it. Firstly, autocompletion can be disabled. Secondly, the most specific constrain can be applied. If they are not compatible in the generality sense, one needs to introduce an autocompletion order, which would resolve the problem by picking only one constraint to suggest the values.

@@

@@

\sergey{ECAI reference style file ignores their guideline and their guideline ignores what is written in the guidelines!}
flashfill, flashextract, flashmeta \cite{flashfill,flashextract,flashmeta}
\begin{itemize}
  \item their supervised vs our unsupervised approach
  \item they look for a single ``smallest'' solution, we enumerate them all
  \item they are looking for a function, we solve constraint satisfaction problems
  \item we do not assume classic row based data layout, we work in the tabular setting
\end{itemize}

>>>>

The question that we answer in this paper is: is it possible to discover or reconstruct structural constraints (relations, functions) in flat tabular spreadsheet data?
To answer this question we contribute a general-purpose method and system, named \sname (pronounced as the word ``tackle''), for discovering row-wise and column-wise constraints.
%Furthermore, the algorithm accepts declarative specifications of the constraints and predicates of interest (the "bias"). \tias{revise declarative claim?}
It operates directly on (headerless) tables of a spreadsheet in an unsupervised setting, as it reasons on that raw tabular data directly with no example constraint instantiations given. %, and discovers column- and row-wise constraints.
We demonstrate the utility of our approach in an experimental evaluation.
Moreover, we sketch additional application scenario's such as autocompletion and error detection.
\\\\
This paper is organized as follows.
Section \ref{sec:formalization} introduces concepts relevant to our approach. Section \ref{sec:approach} presents the problem statement and the approach. Section \ref{sec:evaluation} presents the evaluation of the approach. Section \ref{sec:applications} shows how our system can be used for applications. Section \ref{sec:related_work} presents an overview of the related work. Section \ref{sec:conclusions} provides conclusions.

@@

Our proposed methodology contains the following steps:
\begin{enumerate}
\item Extract headerless tables from tabular data
\item Partition the tables into maximally contiguous, type-consistent \textit{superblocks}
\item Generate for each constraint all valid argument assignments in two steps:
\begin{enumerate}
\item For each constraint $c$, generate all atoms $c(B_1, \ldots ,B_n)$ where each $B_i$ is a maximal \textit{superblock} and the $B_i$ are compatible with the signature.
\item For each generated atom $c(B_1, \ldots ,B_n)$, find all valid $c(B'_1, \ldots, B'_n)$ such that for all $i$ holds $B'_i \sqsubseteq B_i$ and the signature and definition of $c(B'_1, \ldots, B'_n)$ are satisfied.
\end{enumerate}
\end{enumerate}

@@

\begin{figure}[t]
  \centering
  \includegraphics[width=1\linewidth]{figures/tabletool.png}
  \caption{This figure shows the table selection tool of \sname. A user selects rectangular ranges of cells that correspond to (headerless) tables. If the data has an unambiguous orientation, this information can be provided as well.}
  \label{fig:visual_tool}
\end{figure}

@@

\paragraph{Example}
As mentioned before, $\eclookup{G'_2}{G'_1}{H'_1}{H'_2}$ implies $\ecfkey{G'_1}{H'_1}$ and hence the $\textit{LOOKUP}$ constraint template depends on $\textit{FOREIGNKEY}$ in Figure~\ref{fig:learning_order} and will be earlier in the ordering.

@@

\sname supports the ASP \cite{whaisasp} formalism, using the stand-alone Clingo 4.5.4 \cite{clingo} solver, the stand-alone Constraint Programming language MiniZinc \cite{minizinc} with the Gecode 4.4.0 solver which supports floats, as well as an embedded Python-based CP solver \cite{python_constraint}. The constraint signature that must be satisfied to generate the groups can be specified in each of the tree formalisms, to ease prototyping and extendability towards new constraints. The embedded solver is fastest as it does not have the overhead of file-writing and calling an external program, and is used in the experiments below.

@@

% Extending the system with a new constraint template amounts to \tias{TODO: samuel, explain how to define syntax; signature-for-groupgen; signature-for-def; definition}

% \sergey{under question now, might worth deleting} \\
% \sergey{example new template: vr = floor(v), float(v), int(vr), len(v)=len(vr), v,vr are single vectors?}

% \begin{equation}
%   \forall i{:}~ v_r[i] \leq v[i] < v_r[i] + 0.5
%   \label{eq:floor_extension}
% \end{equation}

@@

\subsubsection{Results}
\sname takes a few seconds to find 18 constraints, including all of the 12 solutions described in Figure~\ref{fig:sol_example}.
Of the 6 remaining (redundant) constraints, 5 are $\mathit{rank}$ constraints that are true by accident, such as: \begin{align*}
  & \ecrank{\range{T_1}{\rangeall}{1}}{\range{T_1}{\rangeall}{5}}
\end{align*}
The other redundant constraint is an additional $\mathit{lookup}$ that holds because the two vectors in \range{T_2}{\rangeall}{\rangeto{2}{3}} can both be used to look each other up in \range{T_1}{\rangeall}{\rangeto{1}{2}} and we considered only one of them to be essential (lookup using the ID).

@@

\subsection{Benchmark}
There are three main categories of spreadsheets in the benchmark: spreadsheets from an exercise session teaching Excel at an affiliated University, spreadsheets from tutorials online and publicly available spreadsheets such as crime statistics or financial reports that demonstrate more real world usage\footnote{(attached to to the submitted paper)}.
The case study is also included.

@@

\subsection{Autocompletion}
\begin{figure}[thb]
  \begin{center}
    \includegraphics[width=0.33\textwidth]{figures/autocompletion_example.png}
  \end{center}
  \caption{\sergey{Samuel, Tias have a look at it, should be corrected and extended, like with groups indications and formulas?}}
  \label{fig:autocompletion_example}
\end{figure}
\sergey{I think the time discussion is irrelevant, should it be cut out completely?}

@@

\sergey{there should be a discussion on what to do if multiple constraints satisfy the data and propose predictions, just a sketch here for now}
If multiple constraints satisfy the data and allow autocompletion, then there is a number of ways to handle it. Firstly, autocompletion can be disabled. Secondly, the most specific constrain can be applied. Thirdly, if one is not more specific than the other, one needs to introduce an autocompletion order, which would resolve the problem by picking only one constraint to suggest the values.

@@

\sergey{Either we need to elaborate on the online detection or remove offline-online and simplify the rest, otherwise it is not clear what is going on; again I don't see any point in time here}

@@

flashfill, flashextract, flashmeta \cite{flashfill,flashextract,flashmeta}
\begin{itemize}
  \item their supervised vs our unsupervised approach
  \item they look for a single ``smallest'' solution, we enumerate them all
  \item they are looking for a function, we solve constraint satisfaction problems
  \item we do not assume classic row based data layout, we work in the tabular setting
\end{itemize}
